EDA

### 2024-11-30 other/EDA/EDA.qmd
fixed small and big contracts bar chart

@@

df_filtered <- df %>%
  filter(buyer == "National Defence", amount <= 200000)

# Create the bar plot
barplot(
  height = contract_counts_df$Contract_Count,
  names.arg = contract_counts_df$amount_bin,
  las = 2,  # Rotate x-axis labels to perpendicular
  col = "skyblue",
  border = "black",
  main = "Number of Contracts per 10,000 CAD Award Amount Range",
  xlab = "Award Amount (CAD)",
  ylab = "Number of Contracts",
  cex.names = 0.8  # Adjust label size for readability
)

@@

>>>>

print(num_less_than_100k)

@@

```{r}
# Step 1: Create a new column 'contract_size' categorizing contracts
df <- df %>%
  mutate(
    contract_size = ifelse(amount <= 100000, "Small Contracts (≤ 100k)", "Big Contracts (> 100k)")
  )

# Summarize total amount spent per contract size category
total_spent_df <- df %>%
  group_by(contract_size) %>%
  summarize(
    total_amount_spent = sum(amount, na.rm = TRUE)
  ) %>%
  ungroup()

# Convert the total amount to billions of CAD
total_spent_df <- total_spent_df %>%
  mutate(total_amount_billion = total_amount_spent / 1e9)

# Optional: Display the summarized data
print(total_spent_df)

# Create the bar chart
ggplot(total_spent_df, aes(x = contract_size, y = total_amount_billion, fill = contract_size)) +
  geom_bar(stat = "identity", width = 0.6) +
  geom_text(aes(label = paste0(round(total_amount_billion, 2), "B")), 
            vjust = -0.5, 
            size = 5) +  # Adds labels above the bars
  scale_fill_manual(
    values = c("Small Contracts (≤ 100k)" = "steelblue", "Big Contracts (> 100k)" = "tomato")
  ) +
  scale_y_continuous(
    labels = scales::number_format(scale = 1, suffix = "B"),  # Displays y-axis in billions
    expand = expansion(mult = c(0, 0.1))  # Adds space above the bars for labels
  ) +
  labs(
    title = "Total Amount Spent on Contracts",
    subtitle = "Comparison between Small (≤ 100k CAD) and Big (> 100k CAD) Contracts",
    x = "Contract Size",
    y = "Total Amount Spent (Billions of CAD)"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 18, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5, size = 14),
    axis.text.x = element_text(angle = 0, hjust = 0.5, size = 14),
    axis.text.y = element_text(size = 12),
    axis.title = element_text(size = 16),
    legend.position = "none",  # Removes the legend as it's redundant
    plot.caption = element_text(hjust = 0.5, size = 10, face = "italic")
  )

@@

```{r}
# Load necessary libraries
library(ggplot2)
library(dplyr)
library(scales)

# Assuming 'df' is your data frame and has a column 'amount'

# Step 1: Categorize contracts based on amount
df <- df %>%
  mutate(
    contract_size = ifelse(amount <= 100000, "Small Contracts (≤ 100k)", "Big Contracts (> 100k)")
  )

# Step 2: Ensure the order of contract_size (Small on left, Big on right)
df$contract_size <- factor(df$contract_size, levels = c("Small Contracts (≤ 100k)", "Big Contracts (> 100k)"))

# Step 3: Summarize total amount spent per contract size category
total_spent_df <- df %>%
  group_by(contract_size) %>%
  summarize(
    total_amount_spent = sum(amount, na.rm = TRUE)
  ) %>%
  ungroup()

# Step 4: Convert total amount to billions of CAD
total_spent_df <- total_spent_df %>%
  mutate(total_amount_billion = total_amount_spent / 1e9)

# Optional: Display the summarized data
print(total_spent_df)

# Step 5: Create the ggplot2 bar chart
ggplot(total_spent_df, aes(x = contract_size, y = total_amount_billion, fill = contract_size)) +
  geom_bar(stat = "identity", width = 0.6) +
  geom_text(aes(label = paste0(round(total_amount_billion, 2), "B")), 
            vjust = -0.5, 
            size = 5) +
  scale_fill_manual(
    values = c("Small Contracts (≤ 100k)" = "steelblue", "Big Contracts (> 100k)" = "tomato")
  ) +
  scale_y_continuous(
    labels = scales::number_format(scale = 1, suffix = "B"),  # Y-axis in billions
    expand = expansion(mult = c(0, 0.1))  # Adds space above the bars for labels
  ) +
  labs(
    title = "Total Amount Spent on Contracts",
    subtitle = "Comparison between Small (≤ 100k CAD) and Big (> 100k CAD) Contracts",
    x = "Contract Size",
    y = "Total Amount Spent (Billions of CAD)"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 18, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5, size = 14),
    axis.text.x = element_text(angle = 0, hjust = 0.5, size = 14),
    axis.text.y = element_text(size = 12),
    axis.title = element_text(size = 16),
    legend.position = "none",  # Removes the legend as it's redundant
    plot.caption = element_text(hjust = 0.5, size = 10, face = "italic"),
    panel.grid.major.y = element_line(color = "grey80", linetype = "dashed")  # Adds dashed grid lines
  )
